---
title: "w6_assignment"
author: "Frida Hæstrup"
date: "25/10/2020"
output: html_document
---
Loading packages
```{r}
library(pacman)
p_load(tidyverse, gganimate, gapminder, gifski, png, readr)
```

# 1)
#### Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.

```{r}
# Inspecting data
head(gapminder)

# Defining function for calculating GDP

gdp_function <- function(nation, time){
  this_subset <- gapminder %>%
  subset(country == nation & year == time)
  
  if (!is.numeric(this_subset$pop)) {
    stop("pop must be a numeric vector.")
  }
  if (!is.numeric(this_subset$gdpPercap)) {
    stop("gdpPercap must be a numeric vector.")
  }
  
  this_subset <- this_subset %>% mutate(pop*gdpPercap)
  
  GDP <- this_subset$`pop * gdpPercap`[1]
  
  return(GDP)
}

# Calculating GDP in Denmark during the years 1967, 1977, 1987, 1997, 2007, and 2017
gdp_function('Denmark', 1967)
gdp_function('Denmark', 1977)
gdp_function('Denmark', 1987)
gdp_function('Denmark', 1997)
gdp_function('Denmark', 2007)
gdp_function('Denmark', 2017)

```
In order to calculate the GDP of a nation, we can use the information provided by data set on the population and GDP per capita for each nation per year. By multiplying the 'gdpPercap' column with the 'pop' column, we get the GDP of the nation for that year. This is then put into a function and applied to Denmark for the relevant years.
The results can be seen above. 
  

# 2)
#### Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and print out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70.

```{r}
B_df <- as.data.frame(matrix(0, nrow = 1, ncol = 8))  #new df to be filled with data from loop
colnames(B_df) <- c(colnames(gapminder), "starts_with_B", "range_life_exp")

for (i in gapminder$country){ #looping through countries
  current_subset <- gapminder %>% 
    subset(country == i)
  for (j in current_subset$year){ #looping through years for that country
    new_current_subset <- current_subset %>%
    subset(year == j)
    
    new_current_subset$starts_with_B <- startsWith('B', i) #returns TRUE in new colum if country starts with B 

    #determining size of life exp and adding this info to new column
    if (new_current_subset$lifeExp < 50){ 
      new_current_subset$range_life_exp <- "smaller than 50" 
    }
    if (new_current_subset$lifeExp >= 50 & new_current_subset$lifeExp <= 70){
      new_current_subset$range_life_exp <- "between 50 and 70"
    }
    if (new_current_subset$lifeExp > 70){
      new_current_subset$range_life_exp <- "greater than 70"
    }
    
    B_df <- rbind(B_df, new_current_subset) #binding to B_df for each iteration of the loop
  }
}
head(B_df[,-2])
```
This loop returns a dataframe with information on every country for each year about the range of the lifeExp and whether country starts with 'B'. 

