---
title: "w8_assignment"
author: "Frida Hæstrup"
date: "1/11/2020"
output:
  word_document: default
  html_document: default
---


```{r}
# Loading packages
library(pacman)
pacman::p_load(rvest, dplyr, tidyverse)
```

## What top-rated tv-series can you watch?
When looking for a tv-series to watch, one can go to IMDB to look for the top-rated tv-series of all time. However, often, choosing a tv-series to watch also depends on the possibility of streaming it somewhere online. Here, I make a script that filters away the top-rated tv-series that are not available to stream on Netflix. Thus, when browsing the top-rated list, you do not have to manually go into Netflix to see whether each tv-series is available to watch. Instead, you are provided with a top-rated list only featuring those on Netflix to choose from. 

#### Scraping IMDB
I found a list of the 250 top rated tv-series on IMBD (https://www.imdb.com/chart/toptv/?ref_=nv_tvv_250). Using the Google Chrome extension SelectorGadget, I was able to scrape the titles of the tv-series in this list. This is then loaded into R using the rvest library. 
```{r}
# Specifying url to scrape from
imdb_url <- html("https://www.imdb.com/chart/toptv/?ref_=nv_tvv_250")

# Scraping titles of all tv-series of the url above and adding them to a dataframe
imdb <- as.data.frame(imdb_url %>%
  html_nodes(".titleColumn a") %>% #as specified from the selectorgadet-tool, the ".titleColumn a" tells R that it is the titles that should be extracted
  html_text())

# Changing column title
colnames(imdb) <- "title"
```


#### Scraping Netflix

It seems that Netflix does not have an API that makes it posssible for me to directly scrape the titles of tv-series. However, I found a GitHub-repository with information on all tv-series on Netflix (https://github.com/jameskang410/scraping-netflix/blob/master/TV%20Shows.json). I read this json file into R using the rjson-package. As I am only interested in the titles of the tv-series, I used a loop to extract the titles of all the tv-series and add them to a dataframe. 

```{r}
# Loading rjson package
pacman::p_load(rjson)

# Reading json file with information on tv-series from netflix
json_file <- "https://raw.githubusercontent.com/jameskang410/scraping-netflix/master/TV%20Shows.json"
json_data <- fromJSON(paste(readLines(json_file), collapse="")) #function for reading json-files

# Looping through the data to extract titles of tv-series
netflix <- as.data.frame(matrix(0, nrow = 1, ncol = 1))  #new df to be filled with data from loop

for (i in 1:length(json_data$catalogItems)){ #looping through all tv-series
  current_title <- json_data$catalogItems[[i]]$title #extracting title
    netflix <- rbind(netflix, current_title) #adding title to netflix dataframe
}

# Changing column title
colnames(netflix) <- "title"

```

#### List of available tv-series

Now I have a dataframe ('netflix') with a column indicating the title of all tv-series on netflix and a dataframe ('imdb') with a column indicating the title of 250 top-rated tv-series. In order to check what series from the top-rated list are available to stream on Netflix, I filter out those that do not appear on both lists. 

```{r}
# Filter out those not appearing on both lists
possibilites <- netflix %>% 
  filter(as.factor(title) %in% as.factor(imdb$title))
possibilites
```

From this, we see the 35 out of the 250 top-rated tv-series that are available to watch on Netflix. 
